---
title: "p8105_hw2_yy3296"
author: "Yuze Yuan"
date: "2022-10-04"
output: github_document
---

```{r load_libraries}
library(tidyverse)
library(readxl)
```


### Problem 1

Below we import and clean data from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`. The process begins with data import, updates variable names, and selects the columns that will be used in later parts fo this problem. We update `entry` from `yes` / `no` to a logical variable. As part of data import, we specify that `Route` columns 8-11 should be character for consistency with 1-7.

```{r}
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

As it stands, these data are not "tidy": route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format. This will be useful when focusing on specific routes, but may not be necessary when considering questions that focus on station-level variables. 

The following code chunk selects station name and line, and then uses `distinct()` to obtain all unique combinations. As a result, the number of rows in this dataset is the number of unique stations.

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

The next code chunk is similar, but filters according to ADA compliance as an initial step. This produces a dataframe in which the number of rows is the number of ADA compliant stations. 

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

To compute the proportion of station entrances / exits without vending allow entrance, we first exclude station entrances that do not allow vending. Then, we focus on the `entry` variable -- this logical, so taking the mean will produce the desired proportion (recall that R will coerce logical to numeric in cases like this).

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

Lastly, we write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. As a first step, we tidy the data as alluded to previously; that is, we convert `route` from wide to long format. After this step, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain dataframes with the required stations in rows).

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

### Problem 2

I clean the Mr. Trash Wheel sheet including specifing the sheet in the Excel file and to omit non-data entries, omitting rows that do not include dumpster-specific data, and rounding the number of sports balls to the nearest integer and converts the result to an integer variable.Then, I add a new variable "TrashWheel", and named "MrTrashWheel" which can represent Mr. Trash Wheel dataset in combined dataset. 

```{r}
clean_MrTrashWheel = 
  read_excel("data/Trash Wheel Collection Data (1).xlsx", range = "A2:N549") %>%
  janitor::clean_names() %>%
  mutate(sports_balls = as.integer(round(sports_balls, 0)),
         TrashWheel = "MrTrashWheel")
```

I clean the Professor Trash Wheel sheet including specifing the sheet in the Excel file and to omit non-data entries, omitting rows that do not include dumpster-specific data, and transferring the year variable into character.Then, I add a new variable "TrashWheel", and named "ProTrashWheel" which can represent Professor Trash Wheel dataset in combined dataset. 

```{r}
clean_ProTrashWheel = 
  read_excel("data/Trash Wheel Collection Data (1).xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") %>%
  janitor::clean_names() %>%
  mutate(TrashWheel = "ProTrashWheel",
         year = as.character(year))
```

I combine tidy Mr. Trash Wheel dataset with tidy Professor Trash Wheel dataset. 

```{r}
combine_Mr_Pro = bind_rows(clean_MrTrashWheel, clean_ProTrashWheel)
```

* There are `r nrow(combine_Mr_Pro)` observations, `r ncol(combine_Mr_Pro)` variables which are `r colnames(combine_Mr_Pro)`. 
* The total weight of trash collected by Professor Trash Wheel is `r sum(select(clean_ProTrashWheel, weight_tons))`
* The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(select(filter(select(clean_MrTrashWheel, year, sports_balls), year == 2020), sports_balls))`.

### Problem 3

I clean the data in pols-month.csv and break up the variable mon into integer variables year, month, and day. Then, I replace month number with month name. In addition, I create a president variable taking values gop and dem, and I remove prez_dem, prez_gop, and the day variable by using select function.

```{r}
clean_pols_month = 
  read.csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(month = recode(month, "01" = "January", "02" = "February", "03" = "March", "04" = "April", "05" = "May", "06" = "June", "07" = "July", "08" = "August", "09" = "September", "10" = "October", "11" = "November", "12" = "December"), 
         president = prez_gop,
         president = as.character(president),
         president = recode(president, "0" = "dem", "1" = "gop")) %>%
  select(-day, -prez_dem, -prez_gop)
```
